---
title: "lab04-yowsean-li"
author: "Yowsean Li"
date: "9/21/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# creating a basic data frame
my_table1 <- data.frame(
  col1 = LETTERS[1:5],
  col2 = seq(from = 10, to = 50, by = 10),
  col3 = c(TRUE, TRUE, FALSE, TRUE, FALSE)
)
  
my_table1
```

```{r}
# another way to create a basic data frame
my_list <- list(
  col1 = LETTERS[1:5],
  col2 = seq(from = 10, to = 50, by = 10),
  col3 = c(TRUE, TRUE, FALSE, TRUE, FALSE)
)

my_table2 <- data.frame(my_list)

my_table2
```

```{r}
# strings as strings (not as factors)
my_table3 <- data.frame(
  col1 = LETTERS[1:5],
  col2 = seq(from = 10, to = 50, by = 10),
  col3 = c(TRUE, TRUE, FALSE, TRUE, FALSE),
  stringsAsFactors = FALSE
)

is.data.frame(my_table3)
```

## Starting Lineup of GSW
```{r}
gsw_starting <- data.frame(
  Player = c('Thompson', 'Curry', 'Green', 'Durant', 'Pachulia'),
  Position = c('SG', 'PG', 'PF', 'SF', 'C'),
  Salary = c(16663575, 12112359, 15330435, 26540100, 2898000),
  Points = c(1742, 1999, 776, 1555, 426),
  PPG = c(22.3, 25.3, 10.2, 25.1, 6.1),
  Rookie = c(FALSE, FALSE, FALSE, FALSE, FALSE),
  stringsAsFactors = FALSE
)
gsw_starting
```
```{r}
str(gsw_starting)
```
```{r}
gsw_list <- list(
  Player = c('Thompson', 'Curry', 'Green', 'Durant', 'Pachulia'),
  Position = c('SG', 'PG', 'PF', 'SF', 'C'),
  Salary = c(16663575, 12112359, 15330435, 26540100, 2898000),
  Points = c(1742, 1999, 776, 1555, 426),
  PPG = c(22.3, 25.3, 10.2, 25.1, 6.1),
  Rookie = c(FALSE, FALSE, FALSE, FALSE, FALSE)
)
gsw_starting2 <- data.frame(gsw_list)
gsw_starting2
```
```{r}
gsw_starting$Position <- as.factor(gsw_starting$Position)
gsw_starting$Salary <- as.numeric(gsw_starting$Salary)
gsw_starting$Points <- as.integer(gsw_starting$Points)
gsw_starting$PPG <- as.numeric(gsw_starting$PPG)
gsw_starting
```

```{r}
binded <- cbind(gsw_list)
binded
```

```{r}
dat <- read.csv('nba2017-players.csv', stringsAsFactors = FALSE)
dim(dat)
head(dat)
str(dat, vec.len = 1)
summary(dat)
```

```{r}
tail(dat)
```
### Height < 70 inches
```{r}
dat[dat[,'height'] < 70,]
```
### Centers
```{r}
dat[dat[,'position'] == 'C',c('player','salary')]
```
### Durant
```{r}
durant <- dat[dat$player == 'Kevin Durant',]
durant
```
### UCLA Players
```{r}
ucla <- dat[dat$college == 'University of California, Los Angeles',]
ucla
```
### Rookies
```{r}
rookies <- dat[dat$experience == 0,]
rookies
```
### Rookie Centers
```{r}
rookie_centers <- dat[dat$experience == 0 & dat$position == 'C',]
rookie_centers
```

### Top Players
```{r}
top_players <- dat[dat$games > 50 & dat$minutes > 100,]
top_players
```

### Height Info
```{r}
height_v <- as.vector(dat$height)
max(height_v)
min(height_v)
mean(height_v)
```

### Tallest Players
```{r}
dat[dat$height == max(height_v),1]
```

### Shortest Players
```{r}
dat[dat$height == min(height_v),1]
```

### Unique Teams
```{r}
unique(dat$team)
```
### Number of Teams
```{r}
length(unique(dat$team))
```
### Oldest Player
```{r}
dat[dat$age == max(as.vector(dat$age)),1]
```
### Median Salary of all Players
```{r}
median(dat$salary)
```
### Median salary of players with 10+ yrs of experience
```{r}
median(dat[dat$experience >= 10,'salary'])
```
### Median Salary of SG and PG
```{r}
median(dat[dat$position %in% c('SG','PG'), 'salary'])
```
### Median Salary of PF, 29 Years or older, 74 inches tall or less
```{r}
median(dat[dat$position == 'PF' & dat$age >= 29 & dat$height <= 74, 'salary'])
```
### Number of Players who scored 4 pts or less
```{r}
length(dat[dat$points <= 4,'player'])
```
### Players who scored 4pts or less
```{r}
dat[dat$points <= 4,'player']
```
### Player with 0 points
```{r}
dat[dat$points == 0,'player']
```
### Number of Players from UC Berkeley
```{r}
length(dat[dat$college == 'University of California, Berkeley','player'])
```
### Number of Players from Notre Dame
```{r}
length(dat[dat$college == 'University of Notre Dame', 'player'])
dat[dat$college == 'University of Notre Dame', 'player']
```
### Players with Weight > 260lb
```{r}
length(dat[dat$weight > 260, 'player'])
dat[dat$weight > 260, 'player']
```
### Players who did not attend US College
```{r}
length(dat[dat$college == '', 'player'])
```
### Player with maximum rate of points per minute
```{r}
dat$ppm3 <- NULL
ppm <- dat$points / dat$minutes
dat <-cbind(dat, ppm)
dat[dat$ppm == max(as.vector(dat$ppm)), 'player']
```
### Player with maximum rate of 3 pointers per minute
```{r}
dat$ppm3 <- NULL
ppm3 <- dat$points3 / dat$minutes
dat <-cbind(dat, ppm3)
dat[dat$ppm3 == max(as.vector(dat$ppm3)), 'player']
```
### Player with maximum rate of 2 pointers per minute
```{r}
dat$ppm2 <- NULL
ppm2 <- dat$points2 / dat$minutes
dat <-cbind(dat, ppm2)
dat[dat$ppm2 == max(as.vector(dat$ppm2)), 'player']
```
### Player with maximum rate of 1 pointers per minute
```{r}
dat$ppm1 <- NULL
ppm1 <- dat$points1 / dat$minutes
dat <-cbind(dat, ppm1)
dat[dat$ppm1 == max(as.vector(dat$ppm1)), 'player']
```
### GSW Stats
```{r}
gsw <- dat[dat$team == 'GSW', c('player','height','weight')]
gsw
```
### GSW Sorted by Height
```{r}
gsw[order(gsw$height),]
```
### GSW Sorted by Weight
```{r}
gsw[order(gsw$weight, decreasing = TRUE),]
```

### Top 5 highest paid
```{r}
head(dat[order(dat$salary, decreasing = TRUE), c('player', 'team', 'salary')], n = 5)
```

### Top 10 3 point players
```{r}
head(dat[order(dat$points3, decreasing = TRUE), c('player', 'team', 'points3')], n = 10)
```

### Grouped by Position
```{r}
aggregate(dat[,c('height', 'weight', 'age')], by=list(dat$position), FUN = mean)
```
### Grouped by Team
```{r}
aggregate(dat[,c('height', 'weight', 'age')], by=list(dat$team), FUN = mean)
```
### Grouped by Team and Position
```{r}
aggregate(dat[,c('height', 'weight', 'age')], by=list(dat$position, dat$team), FUN = mean)
```