---
title: "hw01-yowsean-li"
author: "Yowsean Li"
date: "9/14/2017"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# load data file
load("data/nba2017-salary-points.RData")
ls()
```

# 1. Data Preprocessing
```{r}
salary_millions = salary / 1000000
```
```{r}
experience <- as.numeric(replace(experience, experience=="R", 0))
```
```{r}
position_fac <- factor(position)
levels(position_fac) <- c('center', 'power_fwd', 'point_guard', 'small_fwd', 'shoot_guard')
```
```{r}
table(position_fac)
```

# 2. Scatterplot of Points and Salary
```{r}
plot(points, salary_millions, xlab='Points', ylab='Salary (in millions)')
title('Scatter Plot of Points and Salary')
```

# 3. Correlation between Points and Salary
```{r}
n <- length(player)
mean_points <- sum(points) / n
mean_salary <- sum(salary_millions) / n
var_points <- sum((points - mean_points)^2) / (n-1)
var_salary <- sum((salary_millions - mean_salary)^2) / (n-1)
sd_points <- sqrt(var_points)
sd_salary <- sqrt(var_salary)
cov_pts_salary <- sum((points - mean_points)*(salary_millions - mean_salary)) / (n-1)
cor_pts_salary <- cov_pts_salary / (sd_points * sd_salary)

paste('n:', n)
paste('mean_points:', mean_points)
paste('mean_salary:', mean_salary)
paste('var_points:', var_points)
paste('var_salary:', var_salary)
paste('sd_points:', sd_points)
paste('sd_salary:', sd_salary)
paste('cov_pts_salary:', cov_pts_salary)
paste('mean_salary:', mean_salary)
paste('cor_pts_salary:', cor_pts_salary)
```

# 4. Simple Linear Regression
```{r}
b_1 <- cor_pts_salary * (sd_salary / sd_points)
b_0 <- mean_salary - b_1 * mean_points
X <- points
Y <- b_0 + b_1 * X

summary(Y)
```

Summary:

- Minimum: 1.510
- 1st Quartile: 2.845
- Median: 5.206
- Mean: 6.187
- 3rd Quartile: 8.184
- Maximum: 23.298

The regression equation is:
```{r}
Y_hat <- 1.510 + 0.009 * X
```

The slope coefficient b_1 is the average increase in salary per point starting from the intercept.

The intercept term b_0 is the minimum salary.

Predicted Salaries:

- 0 points: 1.510 million
- 100 points: 2.366 million
- 500 points: 5.789 million
- 1000 points: 10.067 million
- 2000 points: 18.623 million

# 5. Plotting the Regression Line
```{r}
plot(points, salary_millions, xlab='Points', ylab='Salary (in millions)')
abline(b_0, b_1)
lines(lowess(points, salary_millions))
text(2300, 17, labels = 'regression')
text(2000, 25, labels = 'lowess')
title('Regression and Lowess Lines')
```

# 6. Regression Residuals and Coefficient of Determination $R^2$
```{r}
residuals <- salary_millions - Y
summary(residuals)
```
```{r}
RSS <- sum(residuals^2)
TSS <- sum((salary_millions - mean_salary)^2)
R2 <- 1 - RSS/TSS
paste('RSS: ', RSS)
paste('TSS: ', TSS)
paste('R2: ', R2)
```
# 7. Exploring Position and Experience
```{r}
plot(experience, salary_millions, xlab='Years of Experience', ylab='Salary (in millions)')
title('Scatter Plot with lowess smooth')
lines(lowess(experience, salary_millions))
```
We can see that salary does increase with more experience, but reaches a peak and slowly starts to decrease at 5 years.

```{r}
library(scatterplot3d)
scatterplot3d(points, experience, salary_millions)
title('3D Scatterplot')
```
This plot shows us that salary does increase as experience and points increase.

```{r}
boxplot(salary_millions~position, xlab="Position", ylab="Salary (millions)")
title('Boxplot of Position and Salary')
```
The boxplot shows that salary does not seem to have a strong correlation with position, although centers seem to be paid the most overall.

# 8. Comments and Reflections
- The hardest parts were figuring out the correct parameters for different functions.
- The easier parts were converting the equations into code.
- This was not my first time using Git or Github.
- I did not need help completing the assignment.
- It took me a couple hours to finish.
- There was no particular part that was most time consuming.
- The most exciting part was drawing the graphs since I could actually visualize the data in different ways.
