---
title: "warmup2"
author: "Yowsean Li"
date: "9/11/2017"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
load("nba2017-salary-points.RData")
ls()
```


###Quantitative Variable - Points

```{r}
summary(points)
sd(points)
```

* Mean: 546.6
* Standard Deviation: 489.0
* Minimum: 0.0
* Maximum: 2558.0
* Median: 432.0
* 1st Quartile: 156.0
* 3rd Quartile: 780.0

```{r}
hist(points)
```

```{r}
boxplot(points)
```

The distribution of points is skewed to the right with a spread from 0 to 2000 points.


### Qualitative Variable - position
```{r}
factor(position)
```

```{r}
table(position)
```

```{r}
table(position)/length(position)
```

We can see that the distribution between positions is very uniform. Each position accounts for around 20% of players.

```{r}
barplot(table(position)/length(position))
```


### Report

* Hard parts were researching R docs to find the right functions to use.
* Easier parts were plotting graphs.
* I didn't encounter any significant errors.
* I still need to learn how to manipulate graphs and data
* The most time consuming part was researching the R docs
* I did not collaborate
* There were no frustrating issues
